
\newacronym{phlt}{PHT}{Probabilistic Hough Line Transform}
\newacronym{hlt}{HLT}{Hough Line Transform}
\newacronym{hot}{\ensuremath{\mathcal{O}(\partial^2)}}{Second order and higher terms}
\newacronym{intensity}{\ensuremath{I_x(x,y,t)}}{Intensity spatial derivatives for $x$ direction at position $(x,y)$ at time $t$}
\newacronym{intensity.spatial}{\ensuremath{I(x,y,t)}}{Intensity at position $(x,y)$ at time $t$}
\newacronym{lvds}{LVDS}{Low Voltage Differential Signalling}
\newacronym{8N1}{8N1}{Common configuration for communicating over a serial port on a computer. Describes the protocol as 8 bits for data, No parity data and 1 stop bit}
\newacronym{rov}{ROV}{Remote Operated Vehicle}
\newacronym{auv}{AUV}{Autonomous Underwater Vehicle}
\newacronym{ttl}{TTL}{Transistor-transistor logic. Describes the voltage levels used for signalling in integrated circuits. Usually 3.3 V or 5 V}
\newacronym{dvl}{DVL}{Doppler Velocity Log}
\newacronym{rxd}{RXD}{Receive Data Pin}
\newacronym{txd}{TXD}{Transmit Data Pin}
\newacronym{gnd}{GND}{Ground Pin}
\newacronym{roi}{ROI}{Region Of Interest}



\newglossaryentry{itk}
{
	name		= {ITK},
	description = {Department of Engineering Cybernetics at the IME Faculty, NTNU}
}

\newglossaryentry{nrk}
{
	name		= {NRK},
	description = {Norwegian Broadcasting Corporation}
}

\newglossaryentry{sintef}
{
	name		= {SINTEF},
	description = {is previously known as Stiftelsen for industriell og teknisk forskning ved Norges tekniske høgskole (NTH), now only SINTEF. The largest independent research organization in Scandinavia}
}

\newglossaryentry{odometric}
{
	name		= {Odometry},
	description	= {is the technique of using data from moving sensors to estimate a change in position of a object over time}
}

\newglossaryentry{dshow}
{
	name		= {DirectShow},
	description = {is a Microsoft\texttrademark technology that is part of the DirectX multimedia framework. 
					It gives a generic interface to operate on media and streams on the Windows platform}
}

\newglossaryentry{wdm}
{
	name		= {Windows Driver Model},
	description = {is the Windows generic framework for device drivers}
}

\newglossaryentry{ocv}
{
	name		= {OpenCV},
	description = {is a free and open source cross-platform library that is tailored towards computer vision. It was originally developed at Intel, but is 
					now available under a BSD license}
}

\newglossaryentry{python}
{
	name		= {Python},
	description = {is a open and dynamically typed high-level programming language. It is mainly designed for fast development and readability}
}

\newglossaryentry{cpp}
{
	name		= {C++},
	description = {is a open and statically typed programming language. Based off the C language, it is known for its good optimization and speed},
	sort		= {cpp}
}

\newglossaryentry{jit}
{
	name		= {JIT},
	description = {stands for Just-in-time compilation. Fragments of code are compiled to machine language on demand}
}

\newglossaryentry{marshalling}
{
	name		= {Marshalling},
	description = {is the process of transcoding or changing the representation of a object in memory for consumption by some receiver. Also known as serialization}
}
